---
title: "ch5,6exercises"
author: "Lauren Temple"
date: "2/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, ggplot2, dplyr, MASS, boot)
```

5.8
a.
```{r}
set.seed(1)
x <- rnorm(100)
y <- x - 2 * x^2 + rnorm(100)
```

n= 100
p=2
Y= X - 2X^2 + e

b.
```{r}
plot(x, y)
```
This plot appears to be quadratic and ranges from -2 to 2 on the x axis and -10 to 2 on the y axis.

c.
```{r}
set.seed(42)
data5.8 = data.frame(x, y)
```

```{r}
i <- glm(y ~ x)
cv.glm(data5.8, i)$delta
```

```{r}
ii <- glm(y ~ poly(x, 2))
cv.glm(data5.8, ii)$delta
```

```{r}
iii <- glm(y ~ poly(x, 3))
cv.glm(data5.8, iii)$delta
```

```{r}
iv <- glm(y ~ poly(x, 4))
cv.glm(data5.8, iv)$delta
```

d.
```{r}
set.seed(123)
data5.8 = data.frame(x, y)
```

```{r}
i <- glm(y ~ x)
cv.glm(data5.8, i)$delta
```

```{r}
ii <- glm(y ~ poly(x, 2))
cv.glm(data5.8, ii)$delta
```

```{r}
iii <- glm(y ~ poly(x, 3))
cv.glm(data5.8, iii)$delta
```

```{r}
iv <- glm(y ~ poly(x, 4))
cv.glm(data5.8, iv)$delta
```

The results are the same with a different seed because LOOCV evaluates n folds of a single observation.

e.
The second model, the quadratic one had the lowest LOOCV test error rate due to the fact that it most closely matches the true form of Y we saw in the scatter plot. 

f.
```{r}
summary(i)
```

```{r}
summary(ii)
```

```{r}
summary(iii)
```

```{r}
summary(iv)
```

The p-values show statistical significance for the linear and quadratic terms but not the terms of higher degree polynomials. 

6.2
```{r}

```

6.9
```{r}

```

6.10
```{r}

```

6.11
```{r}

```

